MISUSES AND INACCURACIES OF PAGEL'S LAMBDA ML ESTIMATION METHODS

### 2020-03-30 Manuscript/Manuscript.tex
All tree simulations completed except the 1028, still running on the lab server.
Rough intro of manuscript developed slightly.

\section{A multivariate equivalency and general test for Rensch's Rule:
Implications for the evolution of sexual shape
dimorphism}\label{a-multivariate-equivalency-and-general-test-for-renschs-rule-implications-for-the-evolution-of-sexual-shape-dimorphism}

@@

\textbf{Short Title}: Problems with Pagel's lambda \hfill\break

@@

Rensch's rule, the allometric association between the degree of sexual
dimorphism and overall body size, is an important framework for
understanding the evolution of trends in phenotypic differentiation
between the sexes across taxa. While virtually all recent treatments
have explored the degree of sexual size dimorphism, understanding how
dimorphism in other phenotypic traits evolves is equally relevant.
Indeed, Rensch's original treatment focused not only on size, but also
on dimorphism in relative body proportions, coloration, the presence of
ornaments, etc. Here, we derive a multivariate equivalency for viewing
trends in sexual dimorphism -- relative to overall body size -- across
taxa, and provide a generalized test to determine whether such patterns
are consistent with Rensch's rule. For univariate linear traits such as
body size, our approach yields equivalent results to those from standard
procedures, but our test is also capable of detecting trends in
multivariate datasets like shape. Computer simulations reveal the method
displays appropriate statistical properties, and an empirical example in
Mediterranean lizards illustrates the efficacy of our approach on both
univariate and multivariate phenotypes. Our generalized procedure
substantially extends the toolkit available for investigating
macroevolutionary patterns of sexual dimorphism and seeking a better
understanding of the processes that underlie them.

@@

It is widely observed that males and females of the same species are not
identical, but rather differ -- sometimes substantially -- in their
phenotypic characteristics (Fairbairn 1997, 2013). Indeed, sexual
dimorphism is pervasive across many animal clades, and understanding the
causes for these patterns has attracted considerable attention in
evolutionary biology for well over a century (e.g., Darwin 1871;
Fairbairn et al. 2007). One of the most conspicuous differences between
males and females is in their body sizes (Andersson 1994), where
sex-specific differences are often hypothesized to be the result of
selection pressures related to reproduction, or the distinct ecological
roles of males and females in their respective habitat (Cox et al. 2003;
Stephens and Wiens 2009; Kaliontzopoulou et al. 2015; Tarr et al. 2018;
Littleford-Colquhoun et al. 2019). For example, both sexual selection on
male body size (Cox et al. 2003; Garcia-Navas et al. 2015; Horne et al.
2020), and fecundity selection on female body size (Serrano-Meneses and
Szekely 2006; Stuart-Fox 2008), are expected to enhance the degree of
sexual size dimorphism in taxa over evolutionary time. Likewise, when
males and females respond differently to environmental factors,
intraspecific competition may generate ecological divergence between
them, resulting in varying patterns of sexual dimorphism (Butler et al.
2000; Bolnick and Doebeli 2003; Dayan and Simberloff 2005;
Kaliontzopoulou et al. 2010; Meiri et al. 2014). \hfill\break

At macroevolutionary scales, changes in the relative contribution of the
selective forces above may create variation in the degree of sexual
dimorphism displayed across closely related taxa. For instance, sister
species inhabiting different environments, or possessing distinct
ecological adaptations, may exhibit differing degrees of sexual
dimorphism, as a result of the varying intensity of sexual selection
they endure (Östman and Stuart-Fox 2011). Likewise, the interplay
between sexual and natural selection that species experience may result
in sexual dimorphism of differing magnitudes across taxa (Blanckenhorn
et al. 2006; Stuart-Fox and Moussalli 2007; Kaliontzopoulou et al.
2015). One evolutionary framework that connects adaptive explanations
for the evolution of sexual size dimorphism (SSD) with trends in overall
body size across taxa is Rensch's rule. Here it is predicted that SSD
will increase with species' average body size in species where males are
the larger sex (hyperallometry), and likewise will decrease with
increasing body size where females are the larger sex (hypoallometry)
(Abouheif and Fairbairn 1997; Fairbairn 1997). Put another way, Rensch's
rule predicts that the extent of sexual dimorphism increases at more
extreme body sizes (both large and small) across taxa. Empirical studies
have documented trends in SSD consistent with Rensch's rule in a wide
variety of animal taxa (Dale et al. 2007; Ceballos et al. 2013; Regis
and Meik 2017; Johnson et al. 2017). However, other clades display the
converse trend (Burbrink and Futterman 2019; Peñalver-Alcázar et al.
2019), and some clades display no relationship between SSD and body size
(Astúa 2010; Hirst and Kiørboe 2014; Johnson et al. 2017). From an
evolutionary perspective, the association between SSD and body size has
been hypothesized to be driven by variation in ontogenetic or static
allometries across taxa, which results from the influence of sexual
selection on growth and maturation patterns (Rensch 1960; Dale et al.
2007). \hfill\break

Interestingly, virtually all evaluations of Rensch's rule have
investigated patterns of sexual size dimorphism, yet Rensch's original
contributions were concerned with a much wider array of phenotypic
traits that can differ between the sexes. These included not only sexual
differences in body size, but also in general body proportions (e.g.,
relative wing and limb lengths), the degree of complexity of different
body structures, the presence of sex-specific ornamentation (e.g.,
horns, body coloration, etc.), and other sex-specific differences
(Rensch 1950, 1960). In like manner, some studies have documented
patterns of sexual shape dimorphism (e.g., Berns and Adams 2013; Kelly
et al. 2013; Kaliontzopoulou et al. 2015), sex-specific differences in
ornamentation (Geist and Bayer 1988; Emlen 2008; Watson and Simmons
2010), or coloration (e.g., Endler 1983). However, such studies have
typically been restricted to one or a few taxa, and have not explained
macroevolutionary variation in sexual dimorphism in a manner that
relates directly to Rensch's rule. Thus, when viewed from this
perspective, Rensch's (1950) vision was far more synthetic, as he sought
to explain the extent to which sexual dimorphism -- across a general
suite of phenotypic characteristics -- scaled with overall body size in
organisms, and why some species displayed a greater degree of sexual
dimorphism than did others. Remarkably, patterns of intraspecific static
allometry in sexually selected traits establishes a direct link between
variation in those traits relative to size, and the evolutionary
allometry across taxa in the degree of sexual dimorphism that may be
expected (see also Rensch 1960; Reiss 1986; Bonduriansky 2007). Under
this framework, the ``exaggeration'' of body structures or other
phenotypic traits with increasing body size is predicted to translate,
at macroevolutionary scales, into increased degrees of sexual dimorphism
with increasing body size. However, despite this clear prediction, and
Rensch's original observations, the evolutionary scaling of the degree
of sexual dimorphism in phenotypic traits other than size, remains
largely unexplored. \hfill\break

Another current limitation in our understanding of Rensch's rule is the
fact that to date, all studies have been conducted exclusively on
univariate data, primarily regarding body size. However, understanding
phenotypic evolution is inherently a multivariate endeavor (Blows 2007;
Collyer et al. 2015; Adams and Collyer 2018, 2019), as processes such as
natural and sexual selection simultaneously operate on more than one
trait (Lande 1979; Lande and Arnold 1983), and can affect variation in
complex multidimensional traits such as shape. To interrogate such
patterns, the analytics used to evaluate evolutionary trends in
phenotypic datasets must be sufficiently general to accommodate this
empirical fact. Unfortunately, all current analytical approaches for
evaluating trends in sexual dimorphism as they relate to overall body
size are explicitly univariate; thereby excluding the possibility of
evaluating patterns in complex multidimensional traits such as
organismal shape. Though is has been extensively documented in sexual
size dimorphism, it remains unknown whether Rensch's rule applies
equally to sexual shape dimorphism, as determining this requires an
analytical framework capable of evaluating trends in sexual dimorphism
in both univariate and multivariate datasets. \hfill\break

In this paper, we derive a multivariate equivalency for inspecting
variation in sexual dimorphism -- relative to overall body size --
across taxa, and provide a generalized test to determine whether they
are consistent with Rensch's rule. Our procedure is appropriate for
evaluating trends in univariate traits measured on a linear scale (such
as body size or relative limb length), as well as in multivariate
datasets representing complex phenotypes (such as sets of body
proportions or landmark-based shape data). For univariate data, the
approach yields equivalent results to what is normally accomplished for
size-based SSD evaluations. However, the approach extends these
procedures for the examination of sexual dimorphism in multivariate
data, thereby increasing the potential for our conceptual understanding
of the consequences of sexual and natural selection on complex
organismal phenotypes. We conduct a series of computer simulations
demonstrating that the approach displays appropriate statistical
properties when evaluating trends in both univariate and multivariate
data. We also provide an empirical example in Mediterranean green
lizards which illustrates the efficacy of our approach on both
univariate and multivariate datasets. Finally, we highlight the new
avenues of evolutionary research that our generalization facilitates,
and make comments on the prospects for future empirical advances.

@@

\subsection{\texorpdfstring{\emph{Conceptual
Development}}{Conceptual Development}}\label{conceptual-development}

To arrive at a general test for Rensch's rule that can accommodate
multivariate data, we must first derive a series of mathematical
equivalencies based upon how interspecific patterns of size dimorphism
are typically evaluated. For sexual size dimorphism, interspecific
patterns are often inspected using a bivariate plot, where the size of
one sex is plotted against the size of the other sex (e.g., Fairbairn
and Preziosi 1994; Abouheif and Fairbairn 1997; Ceballos et al. 2013).
An idealized example is shown in Fig. 1A, with male size plotted against
female size. The diagonal line represents no size dimorphism; or a 1:1
size ratio between the sexes across the range of body sizes for species
in the clade of interest. In this construction (i.e., with females
plotted on the x-axis), deviations from this line represent differential
trends in size dimorphism across taxa. To evaluate such patterns
statistically, a phylogenetic regression is performed, and in this case,
a slope significantly greater than 1.0 (i.e., \(\beta_{1}>1.0\)) would
be treated as evidence that the evolutionary changes in sexual size
dimorphism across taxa are consistent with Rensch's rule (Fairbairn
1997). \hfill\break

An equivalent representation of patterns of sexual dimorphism may also
be found using ratios (e.g., male size:female size). When plotted
against the average size for each species (Fig. 1B), the same
information is conveyed as is found in the more commonly utilized male -
female plot of Fig. 1A. However, in this representation, the line
designating no sexual dimorphism is horizontal, with a slope of zero
(i.e., \(\beta_{1}=0.0\)) and an intercept of 1.0. As before, a
phylogenetic regression may be performed on these data, and in this
case, if the slope is significantly steeper than zero (i.e.,
\(\beta_{1}>0.0\)), patterns consistent with Rensch's rule are observed.
\hfill\break

A third equivalency is shown in Fig. 1C. Here, the Euclidean distance
between male and female means for each species
(\(D_{SD}=\sqrt{(Y_{male}-Y_{female})^2}\)) is shown relative to the
overall size of the species. This plot is similar to a Bland-Altmann
plot (Altman and Bland 1983), except that the Euclidean distance, rather
than a difference score, is plotted along the y-axis (as distances can
accommodate data of different dimensionality). Under Rensch's rule, this
\emph{sexual dimorphism distance} may become greater as species are
progressively larger (hyperallometry, sensu Abouheif and Fairbairn 1997)
or may become greater as species are progressively smaller
(hypoallometry, sensu Abouheif and Fairbairn 1997). In both cases,
patterns correspond to species displaying progressively more sexual
dimorphism at extreme body sizes; a trend consistent with Rensch's rule.
Note that the inflection point in Fig. 1C represents an intermediate
size where there is no sexual dimorphism for the trait, and corresponds
to where the male-female regression line of Fig. 1A crosses the line
representing no sexual dimorphism. This inflection point may represent
the overall mean size across species in the lineage, or may be some
other value. For instance, in some lineages (e.g., turtles: Ceballos et
al. 2013), nearly all species are female biased, in which case the
inflection point for \(D_{SD}\) vs.~size would be found at the far-right
end of the plot. \hfill\break

Finally, from \(D_{SD}\) a \emph{sexual dimorphism distance index}
(\(SDDI\)) may be constructed as:

\[SDDI=-1*D_{SD}\] \[SDDI=+1*D_{SD}\]

Here, \(D_{SD}\) is multiplied by either a +1 or -1, based on a
comparison of male and female trait values for the species. For
instance, when single-valued traits on a linear scale are examined
(e.g., leg length or wing extent), \(D_{SD}\) is multiplied by -1 for
those species where females display larger values than males, and
remains unchanged for species where males are larger than females. This
logic is analogous to that of Lovich and Gibbons (1992) for their index
of overall body size dimorphism. However, from a geometric perspective,
this procedure is tantamount to determining which species in Fig. 1A are
found below the 1:1 size line (i.e., the line representing no size
dimorphism), and multiplying those species' \(D_{SD}\) by -1.
\hfill\break

Unfortunately, applying the same logic to multidimensional trait data is
more complicated, as sex-specific differences are not always guaranteed
to align with some \emph{a priori} direction in the multivariate
dataspace. Thus, for multivariate datasets, one of two procedures may be
used. First, for multivariate sets of traits that are measured on the
same linear scale (e.g., length, width, and height measures), one could
envision a hyperdimensional version of the plot displayed in Fig. 1A,
but where the set of male traits are plotted against the set of female
traits. In this construction, because the traits are quantified on the
same measured scale, the 1:1 size vector representing no sexual
dimorphism does exist, and emanates from the origin of the space
extending in a positive direction as the trait values increase. To
determine whether male trait values deviate more from this vector than
do female trait values, a partial least squares analysis may be
performed (Rohlf and Corti 2000). Partial least squares (PLS) is a
multivariate association procedure which describes the covariation
between sets of traits (in this case, between the set of male traits and
the set of female traits). Scores along the first PLS axis represent the
maximal covariation between male and female traits projected to a single
dimension, and species whose male and female PLS scores are identical
are those displaying no sexual dimorphism. Thus, deviations from that
pattern represent instances of multivariate sexual dimorphism. By
plotting these scores, one may determine which species display female
scores (on \(PLS_1\)) that are larger than their corresponding male
scores (when negative values on \(PLS_1\) correspond to smaller
individuals; otherwise, the converse comparison is used). The \(D_{SD}\)
for these species may then be multiplied by -1 to obtain the \(SDDI\)
for that species. Note that when PLS is used on univariate data, this
procedure obtains values identical to those from the method described
previously where male and female sizes are compared directly. Thus, for
multivariate traits measured on a linear scale, this is a direct
generalization of the standard univariate procedure (see empirical
example below). \hfill\break

Alternatively, one might be interested in examining multidimensional
traits other than linear biometric measurements, such as color or shape.
For instance, one may characterize the shape of males and females using
geometric morphometric methods (Bookstein 1991; Mitteroecker and Gunz
2009; Adams et al. 2013), where the locations of anatomical landmarks
and semilandmarks are used to obtain a set of shape variables. Geometric
morphometric methods also result in a multidimensional characterization
of phenotypic traits, from which the degree of sexual dimorphism for
each species may be quantified using \(D_{SD}\). However, because the
resulting shape variables do not have a one to one correspondence with
linear size measures, there is no \emph{a priori} direction in the
multivariate shape space that is equivalent to the 1:1 size vector found
with sets of linear traits. In fact, for the case of isometry (i.e.,
shape does not change as size changes), all observations are represented
by a single location in the shape space. Thus, while a PLS of male
shapes versus female shapes will describe the covariation between them,
this direction is not guaranteed to align with size, because there is no
direction in the shape space that may be used to unambiguously determine
whether dimorphism is male- or female-biased, as for shape variables
there is no \emph{a priori} directionality. Nonetheless, one may still
quantitatively determine whether the degree of sexual shape dimorphism
covaries with overall body size, by using the sexual dimorphism index as
defined by the dimorphism distance (i.e., \(SDDI=D_{SD}\)) to represent
the degree of multivariate shape dimorphism for each species.
\hfill\break

Once the sexual dimorphism distance index (\(SDDI\)) is calculated, it
may be evaluated against the average size of each species. Here, data
conforming to Rensch's rule will result in significant trends of sexual
dimorphism relative to mean body size (Fig. 1D). As with size ratios,
the line of no sexual dimorphism is a horizontal line, but this time it
is centered on zero, as a distance of zero corresponds to no sexual
dimorphism. Thus, a phylogenetic regression of these data whose slope is
significantly different from zero (e.g., \(\beta_{1}\neq0.0\)) provides
evidence that the evolution of sexual dimorphism is consistent with
patterns expected under Rensch's rule. For cases where traits are
measured on a linear scale, patterns expected under Rensch's rule
exhibit a slope significantly greater than zero (\(\beta_{1}>0.0\)), as
in the usual formulation. Likewise, when using geometric morphometric
shape data (or any other type of mean-centered multidimensional data),
hyperallometric patterns consistent with Rensch's rule would also
display a slope significantly greater than zero (\(\beta_{1}>0.0\)),
indicating that the degree of sexual shape dimorphism increases with
increasing body size. Conversely however, hypoallometric patterns would
display the opposite pattern (\(\beta_{1}<0.0\)), indicating that
smaller species display greater sexual shape dimorphism. \hfill\break

\subsection{\texorpdfstring{\emph{Simulations}}{Simulations}}\label{simulations}

To ascertain the statistical performance of the approach proposed here,
we conducted a series of computer simulations. We evaluated the ability
of SDDI to characterize patterns of sexual dimorphism in both univariate
and multivariate datasets. For each simulation, we generated a
pure-birth phylogeny containing 50 species, on which we simulated
datasets under a Brownian motion model of evolution. For univariate data
(\(p=1\)), a single trait was evolved along the phylogeny, and
represented the female trait values (\(x\)). Next, male values (\(y\))
were simulated in such a manner as to generate a known male:female
relationship (plus random error), using the function:
\(y=\alpha{x}+\mathcal{N}(\mu=0,\sigma=0.1)\). For type I error
simulations, no sexual dimorphism was generated, meaning that the slope,
\(\alpha\), characterized a 1:1 relationship between male and female
trait values (i.e., \(\alpha=1.0\)). For power simulations, we simulated
male values with a slope greater than 1.0 to correspond with patterns
following Rensch's rule (i.e., \(\alpha=1.1\)). A total of 1,000
datasets were generated under each simulation condition. We then
evaluated all datasets using standard approaches (i.e., a phylogenetic
regression of male values versus female values), as well as a
phylogenetic regression of SDDI versus average body size. The proportion
of significant datasets was treated as the type I error (\(\alpha=1.0\))
or power (\(\alpha=1.1\)) respectively. \hfill\break

For multivariate data (\(p=5\)), a slightly different procedure was
utilized. First, average body size (\(s\)) was evolved along the
phylogeny under Brownian motion. Next, female traits (\(\mathbf{X}\))
were simulated on the phylogeny under Brownian motion, using a
\(p\times{p}\) trait covariance matrix with correlations between traits
of (\(r=0.7\)). For simulations where there was no pattern of sexual
dimorphism across species (i.e., type I error simulations) male trait
values were then generated by incorporating random error to the female
trait values, in a manner analogous to that achieved in the univariate
simulations:
\(\mathbf{Y}=\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1\)). For
simulations where patterns of sexual dimorphism covaried with body size
(i.e., power simulations), male trait values were generated by
incorporating random error to the female trait values, but with the
addition that female trait values were pre-multiplied by a scalar value
of body size:
\(\mathbf{Y}=(0.1*s)*\mathbf{X}_{F}+\mathcal{N}(\mu=0,\sigma=0.1\)).
This ensured that male and female trait values covaried with one another
as expected, and also covaried with size in the desired direction. As
before, a total of 1,000 datasets were generated under each simulation
condition. We then evaluated all datasets using a phylogenetic
regression of SDDI versus average body size. The proportion of
significant datasets was treated as the type I error (\(\alpha=1.0\)) or
power (\(\alpha=1.1\)) respectively. \hfill\break

\emph{Simulation Results}: For the univariate dataset, we found that
both a phylogenetic regression of male size on female size, as well as a
phylogenetic regression of \(SDDI\) on body size, resulted in
appropriate type I error rates near the nominal \(\alpha\approx0.05\)
(Fig. 2). Further, when data were simulated under conditions consistent
with Rensch's rule, both procedures were capable of detecting the signal
of the pattern, and did so with very high power. Thus, using the
\(SDDI\) approach proposed here was equivalent to what is normally
utilized for univariate datasets. Likewise, for multivariate data, the
\(SDDI\) approach displayed appropriate type I error
(\(\alpha\approx0.05\)) and high statistical power under the conditions
simulated (Fig. 2). Additionally, multiplying \(D_{SD}\) by \(\pm1\) had
little effect on the outcome for multivariate data, as in both cases
type I error and power results were virtually identical. Thus, methods
for evaluating sexual dimorphism using geometric morphometric data are
anticipated to remain robust. Overall results from these simulations
imply that using the sexual dimorphism distance index (\(SDDI\))
proposed here provides an appropriate means of characterizing patterns
of sexual dimorphism in both univariate and multivariate traits, and
that the procedure is capable of detecting patterns consistent with
Rensch's rule when they are present in phenotypic datasets. \hfill\break

\subsection{\texorpdfstring{\emph{Empirical
Example}}{Empirical Example}}\label{empirical-example}

To illustrate the utility of the approach presented here, we conducted
an analysis of both size and shape dimorphism in a group of
Mediterranean lizards with extensive body size variation. Our dataset
consisted of three data types: male and female values of body size
(snout-vent length: SVL), male and female values for a series of head
measurements (head length, head width, head height, pileus length, and
mouth opening: Fig. 3A), and male and female values of head shape (Fig.
3B). A total of 374 specimens from 21 lineages of green lizards, a
monophyletic clade that includes two genera (\emph{Lacerta} and
\emph{Timon}), were obtained from natural history museum collections.
For each lineage, we obtained samples distributed across the entire
distribution range, resulting in an initial database of 1062 specimens.
In order to take sampling bias into account, we selected the 10 largest
individuals per sex, when sufficient specimens were available, for use
in subsequent size and shape analyses. A detailed account of the
specimens used for this study is available in the Supplemental Material.
\hfill\break

From these specimens, we calculated mean body size for each sex and
lineage. Likewise, we calculated the mean trait values for each sex and
lineage for the set of linear measures. These values were then
standardized by SVL; resulting in a set of head measurements
proportional to body size for each sex \(\times\) lineage combination.
Additionally, we quantified lateral head shape variation using geometric
morphometric methods (Adams et al. 2013). This was accomplished by first
digitizing 17 landmarks and semilandmarks from images of the right side
of the head (Fig. 3B). The thin-plate spline was used to estimate the
position of missing landmarks (Gunz et al. 2009), and the position of
the jaw relative to the skull was standardized using the fixed angle
method (Adams 1999). Next, a generalized Procrustes analysis (Rohlf and
Slice 1990) was performed to eliminate non-shape variation from the
landmark data and align the specimens to a common coordinate system. The
position of semilandmarks (i.e.~11-16, Fig. 3B) was optimized by
minimizing the bending energy (Bookstein 1997). From the aligned
Procrustes coordinates the mean head shape for males and for females was
then calculated for each lineage. Finally, the phylogenetic
relationships among lineages were estimated using published sequences on
five mitochondrial and four nuclear gene fragments, totaling 8383 bp.
These were then used following standard phylogenetic inference
procedures to generate a dated consensus phylogeny for the lineages of
interest (for details see Supplemental Material). \hfill\break

Using these data, we conducted a series of analyses to investigate
whether patterns of sexual size dimorphism (SSD) and sexual shape
dimorphism (SShD) in this group corresponded to expectations under
Rensch's rule. First, we examined patterns of SSD using a phylogenetic
regression of male size against female size. Here, a slope significantly
greater than 1.0 (\(\beta>1.0\)) was treated as evidence consistent with
Rensch's rule. Additionally, we evaluated SSD patterns using the
approach developed here, via a phylogenetic regression of the sexual
size dimorphism index against mean body size for each species. Second,
we evaluated multivariate patterns of sexual dimorphism in head
proportions in relation to body size, using the sexual dimorphism index
obtained from the Euclidean distance between male and female means from
each species (\(D_{SD}\)). Partial least squares was used to determine
which species displayed female scores (on \(PLS_1\)) larger than their
corresponding male scores, and the sexual dimorphism index was then
obtained by multiplying the sexual dimorphism distance by +1 or -1,
respectively (i.e., \(SDDI=\pm1*D_{SD}\)). We then performed a
phylogenetic regression of \(SDDI\) for head proportions versus mean
body size to determine whether patterns of dimorphism corresponded to
what is expected under Rensch's rule. Third, we obtained the degree of
sexual dimorphism in head shape for each species using the Procrustes
distance between male and female means, and from this calculated the
sexual shape dimorphism index (\(SShDI\)) in their original units as
described above (i.e., \(SShDI=D_{SD}\)). We then performed a
phylogenetic regression of sexual shape dimorphism (\(SShDI\)) against
the mean body size for each species to determine whether patterns of
head shape dimorphism were consistent with predictions from Rensch's
rule. Additionally, patterns of head shape dimorphism were further
investigated using principal components analysis of the set of mean male
and female head shapes across species. Finally, we used phylogenetic
correlation (phylogenetic partial least squares: sensu Adams and Felice
2014) to determine whether the degree of head shape dimorphism was
associated with levels of size dimorphism. All analyses were performed
in R 3.6.2 (R Core Team 2019) using the packages \texttt{geomorph}
(Adams and Otárola-Castillo 2013; Adams et al. 2019), \texttt{RRPP}
(Collyer and Adams 2018), and routines written by the authors.
\hfill\break

\emph{Empirical Results}: Using phylogenetic regression, we found a
significant relationship between male size and female size, with a slope
significantly greater than one (\(\beta = 1.39\), \(T=2.91\),
\(P = 0.009\): Fig. 3C). This indicated that the degree of sexual size
dimorphism differed systematically with overall body size; a pattern
consistent with that expected under Rensch's rule. In this case, both
hypoallometry and hyperallometry (\emph{sensu} Abouheif and Fairbairn
1997) were displayed (Fig. 3C). Specifically, in male-biased species,
the degree of sexual dimorphism was greater with increasing body size,
and in female-biased species the degree of sexual dimorphism increased
in smaller species (though the pattern was more pronounced in
male-biased species). Similarly, when using the sexual size dimorphism
index (SSDI), we found a significant and positive evolutionary
relationship between SSDI and mean body size, confirming the Rensch's
rule pattern identified using the commonly-used procedure
(\(\beta = 0.43\), \(F=22.95\), \(Z=1.89\), \(P = 0.004\): Fig. 3D).
\hfill\break

When multivariate patterns of sexual dimorphism were evaluated, we found
that the degree of dimorphism in head proportions increased as body size
increased: a pattern consistent with Rensch's rule (\(F=7.20\),
\(Z=1.41\), \(P = 0.022\): Fig. 3E). Likewise, when patterns of head
shape dimorphism were examined, we found that the degree of sexual shape
dimorphism (SShDI) increased in larger species, and a phylogenetic
regression of sexual shape dimorphism versus mean body size was found to
be significant (\(F=6.05\), \(Z=1.35\), \(P = 0.03\): Fig. 3F). Thus,
for both multivariate datasets, the degree of sexual dimorphism was
enhanced in larger species (hyperallometry; sensu Abouheif and Fairbairn
1997): a pattern consistent with what was expected under Rensch's rule.
To our knowledge, this is the first empirical demonstration of a pattern
following Rensch's rule in multivariate data. \hfill\break

Visualizing patterns of head shape dimorphism using principal components
analysis (Fig. 4), it was evident that males and females occupied
distinct regions of morphospace. Further, the head shapes of males and
females of the smallest lineage, \emph{Lacerta agilis bosnica}, were
distinct from those of the largest species, \emph{Timon lepidus}. When
connected, the sexual shape dimorphism vectors for these two species did
not coincide with one another in the ordination, implying a difference
in sexual shape dimorphism across species. Specifically, the vector for
\emph{T. lepidus} was longer than that for \emph{L. agilis bosnica},
which reflected the greater degree of sexual shape dimorphism exhibited
by this species. Thin-plate spline deformation grids facilitated a
graphical description of these shape differences, where sexual shape
dimorphism in \emph{L. agilis bosnica} was best described as a contrast
between longer-headed males, with an amplified tympanic area; and
short-headed females, with a relatively more reduced tympanic region
(Fig. 4, left). By contrast, sexual shape dimorphism in \emph{T.
lepidus} was both more prominent in the tympanic area, similarly to what
was observed in \emph{L. agilis bosnica}; and also included a
modification of the relative lower jaw area. Indeed, males of \emph{T.
lepidus} were characterized by both a longer posterior head region and
by deeper lower jaw configurations as compared to females of the same
species. \hfill\break

Finally, across taxa, the degree of body size and head shape dimorphism
were significantly correlated with one another (\(r=0.73\),
\(p=0.0001\)), but not when evaluated in a phylogenetic context
(\(r_{PPLS} = 0.361\), \(Z=1.28\), \(P = 0.12\)). Part of this
difference may be explained by the fact that the largest species
(i.e.~members of the genus \emph{Timon}) - which displayed the greatest
degree of both size and shape dimorphism - form a clade within the
broader phylogeny for the group (Fig. 5). Thus, despite the strong
association of size and shape dimorphism, there were fewer changes of
these patterns across branches of the phylogeny. Generally speaking, the
degree of size dimorphism and the degree of shape dimorphism increased
in larger taxa, and heat maps of the evolution of size and shape
dimorphism across the phylogeny indicated a strong association between
the two; particularly in large taxa (Fig. 5). Thus, in this system, it
appeared that evolutionary changes in size between the sexes were
accompanied by concomitant changes in head shape.

@@

A longstanding question in evolutionary biology is understanding why
sexual dimorphism, the degree of differentiation between the sexes,
varies across species. A frequently used schema for exploring
macroevolutionary patterns of sexual dimorphism is Rensch's rule, which
predicts that sexual dimorphism should increase with body size in
species where males are the larger sex, and decrease with increasing
body size in species where females are the larger sex. Remarkably,
virtually all recent studies have focused on macroevolutionary variation
in sexual \emph{size} dimorphism, despite the relevance of sexual
dimorphism in other traits, as encompassed by Rensch's original
treatment (Rensch 1950). This gap has been driven, in part, by the lack
of adequate tools for evaluating Rensch's rule in multivariate traits,
such as color, ornament complexity, or shape. In this article, we fill
this procedural gap by providing an analytical equivalency that allows
the generalization of statistical tests of Rensch's rule for both
univariate and multivariate traits. The approach describes the degree of
sexual dimorphism as a distance between male and female trait values in
phenotype space, thereby facilitating its use on either single-valued
traits (e.g., body size), or complex, multivariate traits
(e.g.~organismal shape). Computer simulations confirm that the approach
exhibits appropriate type I error rates and statistical power, thereby
providing an adequate procedure for testing whether empirical data
comply with the predictions of Rensch's rule. The implementation of our
approach on an empirical study of Mediterranean lizards revealed that
patterns of sexual dimorphism comply with Rensch's rule not only for
body size, but also for relative head proportions and head shape,
offering new insights and triggering novel hypotheses about the possible
proximate and evolutionary causes that may underlie such patterns.
\hfill\break

Our empirical example in Mediterranean green lizards exemplifies a case
study investigating variation in sexual dimorphism -- not only in size
-- but also in multivariate phenotypes, like shape, and how this method
can provide useful insights into the proximate and evolutionary causes
underlying such patterns. First, we found that the degree of sexual size
dimorphism increased as taxa became larger (Fig. 3C, D). This pattern is
compliant with Rensch's rule and corresponds with hyperallometry
(Abouheif and Fairbairn 1997), which is to be expected when males are
the larger sex (Fairbairn 1997). In this group males are larger than
females in most lineages (Fig. 3C), as is commonly observed in lacertids
(Braña 1996). Additionally, as in many lizard species, sexual selection
in this group is primarily mediated through territoriality, where
male-male competition selects for a larger male body size (Stamps 1983).
Thus, in this system there is an association between male-biased sexual
size dimorphism and male-male competition. As such, the patterns of size
dimorphism observed in this study are consistent with those expected
under a scenario where sexual selection drives the evolution of larger
male body size across taxa via male-male competition. \hfill\break

>>>>

\section{Pagel's Lambda Estimates are Often
Inaccurate}\label{pagels-lambda-estimates-are-often-inaccurate}

@@

\textbf{Short Title}: Inaccuracies in Pagel's Lambda \hfill\break

@@

conclusion holds: interpreting the regression is not appreciably
different (in terms of slopes and f values)

@@

Investigating macroevolutionary patterns requires a phylogenetic
approach as species are non-independent by nature of their shared
ancestry. Since the first appropriate method was introduced by
Felsenstein (phylogenetic independent contrasts; {[}Felsenstein1987{]}),
dozens of other methods have been developed and applied to increasingly
complex questions in macroevolutionary biology (e.g.
{[}AdamsNason2019{]}). Understanding the degree of phylogenetic signal
present in a dataset is paramount, and identifies the mode with which a
trait has evolved. High measures of phylogenetic signal indicate a
Brownian motion process, whereas lower levels of phylogenetic signal
indicate natural selection or some other evolutionary force has
influenced the traits evolutionary history. \hfill\break

Several approaches to quantify phylogenetic signal exist. For continuous
data, the most common parameters used in the literature include Pagel's
lambda {[}Pagel1997{]} and Blomberg's kappa {[}Blomberg1900{]}. Pagel's
lambda has the advantage of being couched in the likelihood framework
and thus has also been utilized to encorporate phylogenetic signal while
doing phylogenetic regressions and ANOVA. However, the accuracy of the
lambda estiamtion methods have not been fully evaluated, and thus it
remains unknown the degree to which lambda estimates appropriately
represent degree of phylogenetic signal. \hfill\break

An earlier study {[}BoettigerEtAl2012{]}, breifly addressed this topic
by showing how uninformative smaller phylogenies could be using
estimation methods for various parameters. That paper concluded that a
measure of power must be considered when quantifying Pagel's lambda.
Here we take a more comprehensive approach to demonstrate the scenarios
under which estimated lambdas accurately reflect known lambdas as well
as the effect of these at times dubious estimation methods on
significance testing when used in a pgls framework.

@@

\subsection{\texorpdfstring{\emph{Simulated
trait}}{Simulated trait}}\label{simulated-trait}

To assess the accuracy of Pagel's lambda estimations, we simulated
pure-birth phylogenies with known lambdas. We \_\_\_\_ by scaling
simulated phylogeny with the scaling parameter and . We also did this
with other tree shapes (symmetrical and ladder). \hfill\break

\subsection{\texorpdfstring{\emph{Simulated ANOVA and
Regressions}}{Simulated ANOVA and Regressions}}\label{simulated-anova-and-regressions}

To ascertain the statistical performance of pgls \hfill\break

\subsection{\texorpdfstring{\emph{Meta-Analysis of Empirical
Results}}{Meta-Analysis of Empirical Results}}\label{meta-analysis-of-empirical-results}

Despite the urging of Boettiger and colleagues to publish confidence
intervals with all lambda parameter estimates, only 18\% of papers
published in 2019 do so. \hfill\break

All analyses were performed in R 3.6.2 {[}R-Base{]} using the packages
\texttt{geomorph} (Adams and Otárola-Castillo 2013; Adams et al. 2019),
\texttt{RRPP} (Collyer and Adams 2018). \hfill\break

@@

Using the estimated lambda values from pgls are not useful. The
questions of whether or not signal exists is appropriate, but inferring
more from lambda \emph{magnitude} is inappropriate. \hfill\break

More discussion paragraphs
